---
title: <b>Data Memo</b>
subtitle: "Data Science III (STAT 301-3)"
author: "Alex Chang, Jamie Lee, Grace Park, Nomi Tannenbaum, Susan Tran"
date: "April 6th, 2021"
output:
  html_document:
    toc: true
    toc_float: true
    highlight: "pygments"
    theme: cosmo
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## [**Project Timeline**]{style="color: #AC3B61"}

The due dates are April 21st for an exploratory data analysis and June 12th for the completed project. Given these deadlines, we expect to have our data set loaded into R by Wednesday, April 14th. We will start our analysis as soon as possible after that and will begin to create our predictive model by the end of April. That leaves us with about a month and a half to create different models, evaluate their performance, and draw a conclusion about the data set and a model that might be the "best" fit.

## [**The Data Set**]{style="color: #AC3B61"}

We are interested in predicting the winners for the Academy Awards. Since the 2021 Academy Awards are approaching in the end of April, our group would like to see what factors play heavily into an Oscar win for not only this year's event but also for future Academy Award years. From the Kaggle website, there are a few possible dataset choices that are appropriate for our project. The dataset below contains 27 variables and 441 observations, so it is a relatively small dataset. We may also need to join more sources of data together because our group would also like to focus on the role of gender in the film industry. This dataset does not contain any information on that.

```{r message=FALSE, warning=FALSE}
library(tidyverse)
oscars_data <- read_csv("Data/Oscars-demographics-DFE.csv")
```


## [**Potential Research Questions**]{style="color: #AC3B61"}

Which variables predict (or is related to) an Oscar winner?

## [**Possible Challenges**]{style="color: #AC3B61"}

blah blah blah
