---
title: <b>Data Memo</b>
subtitle: "Data Science III (STAT 301-3)"
author: "Alex Chang, Jamie Lee, Grace Park, Nomi Tannenbaum, Susan Tran"
date: "April 6th, 2021"
output:
  html_document:
    toc: true
    toc_float: true
    highlight: "pygments"
    theme: cosmo
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## [**Project Timeline**]{style="color: #AC3B61"}

The due dates are April 21st for an exploratory data analysis and June 12th for the completed project. Given these deadlines, we expect to have our data set loaded into R by Wednesday, April 14th. We will start our analysis as soon as possible after that and will begin to create our predictive model by the end of April. That leaves us with about a month and a half to create different models, evaluate their performance, and draw a conclusion about the data set and a model that might be the "best" fit.

## [**The Data Set**]{style="color: #AC3B61"}

We are interested in predicting the winners for the Academy Awards, popularly known as the Oscars. Since the 2021 Academy Awards are approaching in the end of April, our group would like to see what factors play heavily into an Oscar win, not only for this year's event but also for future Academy Awards.

From the [Kaggle](https://www.kaggle.com/fmejia21/demographics-of-academy-awards-oscars-winners) website, there are a few possible data set choices that are appropriate for our project. The data set below contains 27 variables and 441 observations, which is relatively small.

```{r message=FALSE, warning=FALSE}
library(tidyverse)
oscars_data <- read_csv("Data/Oscars-demographics-DFE.csv")
```

We may also need to join more sources of data together because our group would also like to focus on the role of gender in the film industry. This data set does not contain any information on that, but it could useful to extract statistics from IMDb about an actor's background, gender, popularity, and so forth.

Another potential data set that we could use, also from the [Kaggle](https://www.kaggle.com/vipulgote4/oscars-nominated-movies-from-2000-to-2017) website, contains information about Oscar nominated movies from 2000 to 2017. 

```{r, message=FALSE}
potential_data <- read_csv("Data/BigML_Dataset_5f50a62c2fb31c516d000176.csv")
```

This data set is considerably larger and has a lot more information on different award ceremonies as well (such as BAFTA, Screen Actors Guild, etc). We could potentially look into merging these two data sets. We would still need to scrape data from IMDb because this data set does not contain any information on the specific actors or directors. 

The last potential data set we are considering, also from [Kaggle](https://www.kaggle.com/unanimad/the-oscar-award), is the largest of the three with over 10,000 observations. It contains Oscars data from 1927-2020 and includes both the individual who won the award and the movie for which they won it. Again, if we wanted to look at the demographics of the winners we would need to collect data from IMDb, but otherwise it seems to be a pretty robust option. 

```{r, message=FALSE}
second_potential <- read_csv("Data/the_oscar_award.csv")
```


## [**Potential Research Questions**]{style="color: #AC3B61"}

While it would be interesting to study which variables might *cause* a change in the outcome, we are more interested in which specific variables *predict* (or are related to) an Oscar winner.

-   What is the most important variable for predicting an Oscar winner? Conversely, what is the least important variable?

-   Which variables might be interactive and therefore may predict an Oscar winner more accurately when those interactions are accounted for?

## [**Possible Challenges**]{style="color: #AC3B61"}

Our data set is limited because there are only 441 observations. This will probably be too small for our model building and selection process. In order to grow our data set, we could scrape data from from the Oscars website, which will provide its own challenges. For example, with the 3rd potential data set mentioned earlier, the data set itself does not contain sufficient variables, therefore requiring that additional information be collected separately which may not be feasible due to the scope. Alternatively we could merge our data set with perhaps a more robust one, which would also pose challenges in the gaps of data upon merging.
