---
title: "Executive Summary"
author: "Grace Park, Susan Tran, Nomi Tannenbaum. Jamie Lee, Alex Chang"
date: "6/9/2021"
output:
  html_document:
    toc: true
    toc_float: true
    theme: paper
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE, comment=FALSE)
```

# Introduction

```{r}
# load package 
library(lubridate)
library(tidyverse)
library(naniar)

# load data
load("data/unprocessed/BigML_Dataset_5f50a62c2fb31c516d000176.csv")
```

Our data set was found on the BigML platform and contains IMDb data scrapes on 1,183 films released between 2000 and 2017 including award nominations, ratings, and award wins. We aim to predict whether or not the movie received any Oscar nominations.

# Exploratory Data Analysis

# Conclusion

Wrapping up our analysis, the performance of our model in predicting whether a film received any Oscar nominations was strong. Next steps could include searching for more context into the meaning of some variables by gaining access to a codebook from the original data source, which we did not have as a resource.
