---
title: "Final EDA"
author: "Grace Park, Susan Tran, Nomi Tannenbaum. Jamie Lee, Alex Chang"
date: "4/25/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Dataset

```{r message=FALSE, warning=FALSE}
# load package 
library(tidyverse)

# load data
load("Data/movies.RData")
```

describing dataset and source here... variables of interest: `critics_rating`, `audience_rating`, `top200_box` , `best_actor_win`, `best_actress_win` , and `best_dir_win` ........ to predict `best_pic_nom`.

## Exploratory Data Analysis

```{r}
skimr::skim_without_charts(movies)

movies %>%
  ggplot(aes(x = best_pic_nom)) +
  geom_bar()
```

missingness issues:

```{r}
miss_var_summary(movies)
```

There are no major missingness issues; the top 3 predictors with the most missingness are actor5, actor4, and actor3, respectively, perhaps meaning that some movies just did not have as many main actors.

```{r}
movies %>% 
  ggplot(aes(x = mpaa_rating, fill = best_pic_nom)) +
  geom_bar(position = "dodge")
```

This bar plot shows that the dataset is representative of a lot of R-rated movies; however, 

